---
title: "Homework 1, STATS 315A"
subtitle: "Stanford University, Winter 2019"
author: "Joe Higgins"
header-includes:
   - \usepackage{amsmath}
output: pdf_document
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
# if you don't have a package, you can get it with "install.packages()"
require('MASS')
require('ggplot2')
require('GGally')
require('dplyr')
require('reshape2')
require('locfdr')
require('mvtnorm')
require('car')
require('Boom')
require('bestglm')
require('BMS')
t.test.p <- function(x,y) {
  tst = t.test(x,y)
  return(tst$p.value)
}

```


## Question 1

Linear versus Knn: you are to run a simulation to compare KNN and linear regression in terms of their performance as a classifier, in the presence of an increasing number of noise variables. We will use a binary response variable $Y$ taking values $\{0,1\}$, and initially $X$ in $R^{2}$. The joint distribution for $(Y,X)$ is
$(1-\pi) f_0(x)$ if $y = 0$
$h_{YX}(y,x)= \pi f_1(x)$ if $y = 1$
where $f_0(x)$ and $f_1(x)$ are each a mixture of K Gaussians:
$f_j(x) = \sum_{k=1}^{K} \omega_{kj}\phi(x;\mu_{kj},\Sigma_{kj}), j = 0,1$ (1)
$\phi(x;\mu,\Sigma)$ is the density function for a bivariate Gaussian with mean vector $\mu$ and covariance matrix $\Sigma$, and the $0 < \omega_{kj} < 1$ are the mixing proportions, with  $\sum_k \omega_{kj} = 1$.

(a) We will use $\pi = .5$, $K=6$, $\omega_{kj} = \frac{1}{6}$ and $\Sigma_{kj} = \sigma^2I= 0.2 = I$, $\forall k,j$. The six location vectors in each class are simulated once and then fixed. Use a standard bivariate gaussian with covariance $I$, and mean-vector $(0,1)$ for class 0 and $(1,0)$ for class 1 to generate the 12 location vectors.

```{r}
#initialize Beta
rm(list = ls())
pi <- 0.5
K <- 6
J <- 1
omega <- matrix(1/6, nrow = K, ncol = J)
Sigma <- 0.2 * diag(2)

location_vectors <- c()

ks <- seq(1,K,1)
js <- seq(0,J,1)
for(k in ks){
  for(j in js){
    print(k)
    print(j)
  }
}


```
