---
title: "Homework 3, STATS 315A"
subtitle: "Stanford University, Winter 2019"
author: "Joe Higgins"
header-includes:
   - \usepackage{amsmath}
output: pdf_document
---

```{r setup, include = FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
# if you don't have a package, you can get it with "install.packages()"
require('MASS')
require('ggplot2')
require('GGally')
require('dplyr')
require('reshape2')
require('locfdr')
require('mvtnorm')
require('car')
require('Boom')
require('bestglm')
require('BMS')
require('class')
require('glmnet')
require('readr')
require('caret')
require('spline')
require('e1071')

t.test.p <- function(x,y) {
  tst = t.test(x,y)
  return(tst$p.value)
}

```

## Question 1

Ex. 18.9 Compare the data piling direction of Exercise 18.8 to the direction of the optimal separating hyperplane (Section 4.5.2) qualitatively. Which makes the widest margin, and why? Use a small simulation to demonstrate the difference.

```{r}
rm(list = ls())

#Parameters
p <- 25
n <- 20
reps <- 1000

margins <- matrix(,nrow=reps,ncol=2)
for(i in 1:reps){
  #Create covariates
  X <- matrix(mvrnorm(p*n, 0, 1), nrow=n, ncol=p)
  s <- svd(X)
  U <- s$u
  D <- diag(s$d)
  V <- s$v
  
  #Create response
  y <- matrix(rbinom(n, 1, .5))
  y[y == 0] <- -1
  
  #Create maximal data piling direction
  B_0 <- V %*% ginv(D) %*% t(U) %*% y
  
  #Margin from maximal data piling direction
  DP_margin <- 2/norm(B_0,"2")
  
  #Margin from SVM
  svmfit <- svm(y ~ ., data = X, kernel = "linear", cost=1000, scale = FALSE, type='C-classification')
  
  beta <- drop(t(svmfit$coefs)%*%X[svmfit$index,])
  beta0 <- svmfit$rho
  f_hat <- (X %*% beta - beta0) #same as svmfit$decision.values

  SVM_margin <- 2/norm(beta,"2")
  
  #Save output
  margins[i,1] <- DP_margin
  margins[i,2] <- SVM_margin
}

margin_difference <- margins[,2] - margins[,1]
cat("Number of times DP margin larger than SVM margin: ",
    length(margin_difference[margin_difference < 0])
)
plot(1:reps, margin_difference)

```

## Question 5

You have a set of $p$ variables in the matrix $X$ which represent some economic indicators measured over a number of years, and a response $y$ also measured for the same years. You fit a linear model, but are criticized because you are told that the regime changes in time, and so should your model. You decide to let your regression coefficients (and intercept) change smoothly with time, a so-called varying coefficient model.

(a) Describe how you could achieve this using natural cubic splines.

(b) Read in the data in `vcdata.csv`. There is a two-column $x$ and a time variable $t$. Fit the varying coefficient model, and plot the three coefficient (functions) versus time (on the same plot using matplot).

```{r}
#Setup, read data
rm(list = ls())
data <- read.csv("vcdata.csv")

#User parameters
df <- 5

#Design data
column_names <- c('intcp', 'x1', 'x2')
p <- length(column_names)
n <- nrow(data)

t_splines <- ns(data$t, df = df, intercept = TRUE)
matplot(data$t, t_splines[,])
intcp <- rep(1, nrow(data))
X <- data.frame(matrix(c(intcp, data$x.1, data$x.2), nrow=n, ncol=p))
colnames(X) <- column_names

#Fit model, aggregate parameters, make predictions
model <- lm(data$y ~ t_splines:(intcp + x1 + x2) - 1, data=X)
bint_t <- t_splines[,] %*% model$coefficients[0*df+1:df]
b1_t <-   t_splines[,] %*% model$coefficients[1*df+1:df]
b2_t <-   t_splines[,] %*% model$coefficients[2*df+1:df]
y_hat <- bint_t + (b1_t * data$x.1) + (b2_t * data$x.2)

#Plot aggregate parameters as function of t
coeff_t <- matrix(c(bint_t, b1_t, b2_t), nrow=nrow(data),ncol=p)
matplot(
  data$t, coeff_t,
  type = "l",
  lty = c('solid','solid','solid'),
  lwd = c(2,2,2),
  col = c('black', 'red', 'green')
)
legend(
  'right', 
  legend = c('B_0','B_x1','B_x2'), 
  col = c('black', 'red', 'green'),
  fill = c('black', 'red', 'green')
)
```

(c) Compute the pointwise standard errors for each of these, and in- clude a standard-error band (upper and lower) for each function.

```{r}
#Create confidence intervals for each 
bands <- c()
for(i in 1:ncol(X)){
  #Make synthetic data, all 1s for the coefficient we want to calculate
  synth_X <- matrix(0,nrow=nrow(data),ncol=p)
  synth_X[,i] <- 1
  colnames(synth_X) <- column_names
  
  #Make "predictions" of B_i, with confidence bands
  bands_i <- predict(model, data.frame(synth_X), interval = 'confidence')
  bands <- cbind(bands,bands_i)
}

#Plot confidence bands of each coefficient
matplot(
  data$t, bands,
  type = "l",
  lty = c('solid','dashed','dashed','solid','dashed','dashed','solid','dashed','dashed'),
  lwd = c(2,1,1,2,1,1,2,1,1),
  col = c('black','black','black','red','red','red','green','green','green')
)
```

think of this as standard errors of y_hat, because its like a regression.

(d) Is this an interaction model? If so, what order.

Yes this is an interaction model. It is a second-order interaction model, since the covariates of the model are based on two vectors multiplied element-wise on each other.

